---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Dana Jensen"
date: "August 10, 2017"
output: html_document
---
```{r setup, include=FALSE}
#prelude
knitr::opts_chunk$set(echo = TRUE)
getwd()
setwd("C:/Users/Dana/Desktop/METHODS III/assignmentIpt.III")
library(pacman)
p_load(modelr, lmerTest, Metrics, lme4, merTools, ggplot2, hydroGOF, gstat, dplyr, stringr, plyr, caret)

##Currupt data, so we use kenneth
ken_train= read.csv("df_train.csv")
ken_test = read.csv("df_test_good.csv")
ken_test = select(ken_test, VISIT, SUBJ, Ethnicity, Diagnosis, Age, ADOS, types_CHI, tokens_CHI, CHI_MLU, verbalIQ, nonVerbalIQ, types_MOT, MOT_MLU, tokens_MOT)
```

### Exercise 1) Testing model performance

```{r}
# Compare the results on the training data () and on the test data. Report both of them. Discuss why they are different.
# - recreate the models you chose last time (just write the code again and apply it to Assignment2TrainData1.csv)

#making the model
model = lmer(CHI_MLU ~ 1 + VISIT + verbalIQ + (1+VISIT | SUBJ), ken_train, REML = F)
#let's see the results
summary(model)
#alright, how good is this model?
modelr::rmse(model, ken_train) #0.344

# - test the performance of the models on the test data 
modelr::rmse(model, ken_test)#0.349

# - optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())
predictInterval(model, ken_test)#not sure how to interpret this?

```

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)
```{r}
# - Create the basic model of ChildMLU as a function of Time and Diagnosis (don't forget the random effects!).
#make a model
model1 = lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|SUBJ), ken_train, REML = F)

# - Make a cross-validated version of the model. (Tips: google the function "createFolds";  loop through each fold, train a model on the other folds and test it on the fold)

#make folds
folds <- cut(seq(1,nrow(ken_train)),breaks=4,labels=FALSE)
#Perform 4 fold cross validation
train_error = NULL
test_error = NULL
for(i in 1:4){#i is numbers 1 through 4 inclusive
    #Segement data by fold using the which() function 
    testIndexes <- which(folds==i,arr.ind=TRUE)#folds
    testData <- ken_train[testIndexes, ]#save the testing folds in variable
    trainData <- ken_train[-testIndexes, ]#save train folds in variable 
    model1 = lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|SUBJ), trainData, REML = F)#nice model m8
    train_error[i] = modelr::rmse(model1, trainData)#take a look at the training error (not bad, not bad)
    testData$prediction <- predict(model1, testData, allow.new.levels = TRUE)#and now, predict!
    test_error[i] = Metrics::rmse(testData$CHI_MLU, testData$prediction)#here we see how well the prediction model did on child mlu (not that great, *sad trumpet*)
    #just for curiosity, you can look at estimates to see how they change
    est1 = fixef(model1)[1]
    }

### - Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.
 
#alright, imma just remake the loop and trial and error the models
folds <- cut(seq(1,nrow(ken_train)),breaks=4,labels=FALSE)
train_error = NULL
test_error = NULL
for(i in 1:4){
    testIndexes <- which(folds==i,arr.ind=TRUE)#folds
    testData <- ken_train[testIndexes, ]#save the testing folds in variable
    trainData <- ken_train[-testIndexes, ]#save train folds in variable 
    model2 = lmer(CHI_MLU ~ 1 + VISIT + verbalIQ + (1+VISIT | SUBJ), trainData, REML = F)
    train_error[i] = modelr::rmse(model2, trainData)
    testData$prediction <- predict(model2, testData, allow.new.levels = TRUE)
    test_error[i] = Metrics::rmse(testData$CHI_MLU, testData$prediction)
    
}
print(train_error)
print(test_error)

# #model2 = lmer(CHI_MLU ~ 1 + VISIT + verbalIQ + (1+VISIT | SUBJ), ken_train, REML = F)
# > print(train_error)
# [1] 0.3447835 0.3566611 0.3550076 0.3429060
# > print(test_error)
# [1] 0.3647887 0.3287494 0.3340756 0.3700598
#ooooo fuck not bad eh?
#


# - Bonus Question 1: How would you go comparing the performance of the basic model and the cross-validated model on the testing set?

# - Bonus Question 2: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?

# - Bonus Question 3: compare the cross-validated predictive error against the actual predictive error on the test data
```
### Exercise 3) Assessing the single child

```{r}
# 
# Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.
#make subset for bernie-boy
bernie = subset(ken_test, SUBJ == 2)
# You want to evaluate:
# 
# - how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.

#group for only td kids
td_group = subset(ken_train, Diagnosis == "TD")
#model that shit
model3 = lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ), td_group, REML = F)
#here comes bernie-boy
bern_model = lm(CHI_MLU ~ VISIT, bernie)
#summarize output
summary(model3)
summary(bern_model)

# (Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; ***input the average of the TD group  for each parameter in the model as X1, X2, etc.).*** - WHAT DOES THIS MEAN??
# 
# - how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```
### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!
```{r}
# Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).
 
# Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.
 
# - Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?
```

### OPTIONAL: Exercise 5): Using Lasso for model selection
Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!


