---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Dana Jensen"
date: "August 10, 2017"
output: html_document
---
```{r setup, include=FALSE}
#prelude
knitr::opts_chunk$set(echo = TRUE)
getwd()
setwd("C:/Users/Dana/Desktop/METHODS III/assignmentIpt.III")
library(pacman)
p_load(modelr, lmerTest, Metrics, lme4, merTools, ggplot2, hydroGOF, gstat, dplyr, stringr, plyr, caret)

#Currupt data, so we use kenneth
ken_train= read.csv("df_train.csv")
ken_test = read.csv("df_test_good.csv")
ken_test = select(ken_test, VISIT, SUBJ, Ethnicity, Diagnosis, Age, ADOS, types_CHI, tokens_CHI, CHI_MLU, verbalIQ, nonVerbalIQ, types_MOT, MOT_MLU, tokens_MOT)
ken_test$SUBJ = ken_test$SUBJ + 100
```

### Exercise 1)
```{r}
##Compare the results on the training data () and on the test data. Report both of them. Discuss why they are different.
##recreate the models you chose last time (just write the code again and apply it to Assignment2TrainData1.csv)

#making the model
model = lmer(CHI_MLU ~ 1 + VISIT + verbalIQ + (1+VISIT | SUBJ), ken_train, REML = F)
#let's see the results
summary(model)
#alright, how good is this model?
modelr::rmse(model, ken_train) #0.344

##test the performance of the models on the test data 
Metrics::rmse(ken_test$CHI_MLU,predict(model, ken_test, allow.new.levels = TRUE))
##optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())
predictInterval(model, ken_test)#not sure how to interpret this?

```

### Exercise 2)
```{r}
##Create the basic model of ChildMLU as a function of Time and Diagnosis (don't forget the random effects!).

#make a model
model1 = lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|SUBJ), ken_train, REML = F)

##Make a cross-validated version of the model
model1 = lmer(CHI_MLU ~ VISIT + Diagnosis + (1+VISIT|SUBJ), trainData, REML = F)

#make folds
fold_function = function(train_df, test_df,model,x){
train_error = NULL
test_error = NULL
n = 1
folds <- cut(seq(1,nrow(train_df)),breaks=x,labels=FALSE)
for(i in 1:x){#i is numbers 1 through 4 inclusive
    #Segement data by fold using the which() function 
    testIndexes <- which(folds==n,arr.ind=TRUE)#folds
    testData <- train_df[testIndexes, ]#save the testing folds in variable
    trainData <- train_df[-testIndexes, ]#save train folds in variable
    model = model
    train_error[n] = modelr::rmse(model, trainData)#take a look at the training error (not bad, not bad)
    testData$prediction <- predict(model, testData, allow.new.levels = TRUE)#and now, predict!
    test_error[n] = modelr::rmse(model, testData)#here we see how well the prediction model did on child mlu (not that great, *sad trumpet*)
    n = n+1
    print(train_error)
    print(test_error)
}}
#Perform 4 fold cross validation
fold_function(ken_train, ken_test,model1, 4)


##Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.
 
model2 = lmer(CHI_MLU ~ 1 + VISIT + verbalIQ + (1+VISIT | SUBJ), ken_train, REML = F)
fold_function(ken_train, ken_test, model2, 4)
# > print(train_error)
# [1] 0.3447835 0.3566611 0.3550076 0.3429060
# > print(test_error)
# [1] 0.3647887 0.3287494 0.3340756 0.3700598
#ooooo fuck not bad eh?

##Bonus Question 1: How would you go comparing the performance of the basic model and the cross-validated model on the testing set?

#-> root mean squared error

##Bonus Question 2: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?

#-> as folds go up, reliability goes down
fold_function(ken_train, ken_test,model1,10)

##Bonus Question 3: compare the cross-validated predictive error against the actual predictive error on the test data
Metrics::rmse(ken_test$CHI_MLU,predict(model, ken_test, allow.new.levels = TRUE))#0.67
mean(test_error)#0.479
```
### Exercise 3)

```{r}
##Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

#make subset for bernie-boy
bernie = subset(ken_test, SUBJ == 2)

##how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.

typical_td = data.frame(SUBJ = rep(200, 6), Diagnosis = rep('TD', 6), VISIT = seq(1, 6), verbalIQ = rep(mean(ken_train$verbalIQ[ken_train$Diagnosis == 'TD']), 6)) #making a typical TD child

#group for only td kids
td_group = subset(ken_train, Diagnosis == "TD")
#model that shit
model3 = lmer(CHI_MLU ~ VISIT + (1+VISIT|SUBJ), td_group, REML = F)
predict(model3, typical_td, allow.new.levels = T)
#get absolute difference
abs(predict(model3, typical_td, allow.new.levels = T) - bernie$CHI_MLU)

#here comes bernie-boy
bern_model = lm(CHI_MLU ~ VISIT, bernie)
#summarize output
summary(model3)
summary(bern_model)
fixef(model3)[1]
coef(bern_model)[1]

##(Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; input the average of the TD group  for each parameter in the model as X1, X2, etc.).

##how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```

### OPTIONAL: Exercise 4) 
```{r}
##Model Selection via Information Criteria, another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

##Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).
model4 = lmer(CHI_MLU ~ 1 + VISIT + verbalIQ + (1+VISIT | SUBJ), ken_train, REML = F)
model5 = lmer(CHI_MLU ~ 1 + VISIT + MOT_MLU + (1+VISIT | SUBJ), ken_train, REML = F)
model6 = lmer(CHI_MLU ~ 1 + VISIT + Diagnosis + (1+VISIT | SUBJ), ken_train, REML = F)

##Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.
I_have_a_big_BIC = anova(model4,model5,model6)
#model4 woot-woot!

##Bonus question: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?
models = c(model4, model5, model6)
folds <- cut(seq(1,nrow(ken_train)),breaks=4,labels=FALSE)
train_error = NULL
test_error = NULL
#loop-de-loop
for (m in models){
  for(i in 1:4){
      testIndexes <- which(folds==i,arr.ind=TRUE)#folds
      testData <- ken_train[testIndexes, ]#save the testing folds in variable
      trainData <- ken_train[-testIndexes, ]#save train folds in variable 
      model = m
      train_error[i] = modelr::rmse(model, trainData)
      testData$prediction <- predict(model, testData, allow.new.levels = TRUE)
      test_error[i] = Metrics::rmse(testData$CHI_MLU, testData$prediction)
      print(m)
      print(train_error)
      print(test_error)
      }}
```

### OPTIONAL: Exercise 5)
```{r}
##Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!


